---
title: "R Notebook"
output: html_notebook
---

Today I wanted to walk through a quick example, combining scraping, calls to the Yahoo finance api and simple asset analysis.I also want to focus on functional programming and tidy iteration.  Today I will be using the R programming langauge.  If you have read any of my posts on linkedin or Medium in the past, you may have noticed that I ussually program in python.  In general, I prefer the python programming language because it has simpler synthax, wider adoption and is more versatile.  

In the R programming langauge there is a set of packages that make up what is called the **tidyverse**. These packages are mostly maintained by engineers at Rstudio and provide a simple, integrated and uniform way to maninipulate data in R.  



I'll start by importing a the packages that I'll need. You'll notice that most of the packages that I use are from the tidyverse.

```{r, package import}
# for scraping
library(rvest)
# blanket import for core tidyverse packages
library(tidyverse)
# tidy financial analysis 
library(tidyquant)
# tidy data cleaning functions
library(janitor)
```


The next thing that I will do is define a variable with todays date.  I will then use the subtract 3 months from todays date.  This returns a another date object, indicating what day came 3 months before today.  I will need this because I want to get the last 3 months of OHLC data for each ticker.

```{r}
# save current system date to a variable
today <- Sys.Date()
# subtract 3 months from the current date
date = today %m+% months(-3)
print(date)
```


I'm going to use the tidyquant package to get the financial data for all SP500 tickers.  The tidyqunat packages core function is tq_get() which can be used to get various information about stocks.  If I pass a string containing a ticker name to tq_get(), it will return Open, High, Low, Close or OHLC data.
```{r}
# pass SP500 ticker ^GSPC to tq_get function
one_ticker = tq_get("^GSPC", from = date)
one_ticker %>% 
  head()
```
Theres a few things to note above:

* tq_get() returns a tidy dataframe
* the ticker name is not in the dataframe
* the '%>%' operator is called a pipe. It passes the object that proceeds it as the first argument to the function that follows it.

I want this OHLC data for all SP500 tickers. In order to do this I will need to do a few things:

* Get a list of all SP500 tickers
* Iterate over this list and call tq_get on each element of the list, returning a dataframe for each ticker
* combine all these dataframes into one dataframe 

Wow! That sounds a little complicated right? luckily, with R, going about this will be pretty simple.  Wikiepdia has a table of all 505 SP500 tickers (some companies, like Google, have multiple asset classes) located at this URL:

https://en.wikipedia.org/wiki/List_of_S%26P_500_companies

To get all the SP500 tickers we are going to scrape this table, using the rvest package.  The Rvest package is a simple scraping package in R that is very similar to python's beautiful soup. In programming, scraping is defined as programatically collecting human readable content from the internent and webpages.

In the code below I scrape the wikipedia table and create a list of all SP500 tickers.  The hardest part of scraping is figuring out the xpath or css to indicate which html nodes to select. I really don't know much about html or css, but using Google Chrome I was able to find the correct xpath.  

```{r}
# get the URL for the wikipedia page with all SP500 symbols
url <- "https://en.wikipedia.org/wiki/List_of_S%26P_500_companies"
# use that URL to scrape the SP500 table using rvest
tickers <- url %>%
  # read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="constituents"]') %>% 
  html_table()
#create a list of tickers
sp500tickers <- tickers[[1]]
sp500tickers = sp500tickers %>% mutate(Symbol = case_when(Symbol == "BRK.B" ~ "BRK-B",
                                           Symbol == "BF.B" ~ "BF-B",
                                            TRUE ~ as.character(Symbol)))
```



```{r}
symbols = sp500tickers$Symbol
# define a function that retunrs OHLC data with the ticker name
get_symbols = function(ticker = "AAPL"){
  df = tq_get(ticker, from = date)  %>%  mutate(symbol = rep(ticker, length(date)))
  return(df)
}
#create the dataframe of SP500 data by interating over our list of symbols and call our get symbols function each time
#the map function accomplishes this

tickers_df = map(symbols, get_symbols) %>% bind_rows()
tickers_df = tickers_df %>% 
  # left join with wikipedia data
  left_join(sp500tickers, by = c('symbol' = 'Symbol')) %>% 
  # make names R compatible
  clean_names() %>% 
  # keep only the columns we need
  select(date:security, gics_sector, gics_sub_industry)


#tickers_df = tq_get(symbols)
Rows = dim(tickers_df)[1]
Columns = dim(tickers_df)[2]
dimensions = data.frame(Rows, Columns)
print(dimensions)
```

```{r}
tickers_df %>% head()
```

```{r}
ticker = "GOOGL"
tickers_df %>% 
  filter(symbol == !!ticker) %>% 
  ggplot(aes(date, adjusted))+
  geom_line()
```

```{r}
daily_sector = tickers_df %>% group_by(security, gics_sector, symbol) %>% 
tq_transmute(select     = adjusted, 
              mutate_fun = periodReturn, 
              period     = "daily")
```


```{r}
avg_return =daily_sector %>% 
  group_by(security, gics_sector, symbol) %>%
  summarise(avg_return = round(mean(daily.returns), 4),Volatility = sd(daily.returns), cv = Volatility/avg_return) %>%         arrange(desc(avg_return), desc(Volatility))
avg_return %>% head()
```

```{r}

avg_return %>% head(20) %>% ggplot(aes(reorder(security, -avg_return), avg_return, fill = avg_return))+
  geom_col()+
  coord_flip()+
  labs(title = "Assets With Highest Returns In SP500 Over Past 3 Month", x = "Security", y = "Average Return")+
  theme_classic()+
  theme(legend.position="none")
```


```{r}
tickers_df %>% 
  filter(security == 'Dow Inc.') %>% 
  ggplot(aes(date, adjusted))+
  geom_line()
```
```{r}
hardware = c('Intel Corp.', 'Advanced Micro Devices Inc', 'Nvidia Corporation')
tickers_df %>% 
  filter(security %in% hardware) %>% 
  select(date, adjusted, security) %>% 
  ggplot(aes(date, adjusted))+
  geom_point()+
  geom_smooth()+
  theme_classic()+
  facet_wrap(~security, scales = 'free_y')
```

```{r}
daily_sector %>% 
  filter(security %in% hardware) %>% 
  ggplot(aes(daily.returns))+
  geom_density(fill = 'grey')+
  facet_wrap(~security)+
  theme_classic()
```

```{r}
avg_return %>% tail(20) %>% ggplot(aes(reorder(security, avg_return), avg_return, fill = avg_return))+
  geom_col()+
  coord_flip()+
  labs(title = "Assets With Lowest Returns In SP500 Over Past 3 Month", x = "Security", y = "Average Return")+
  theme_classic()+
  theme(legend.position="none")
```
```{r}
daily_sector %>% group_by(gics_sector) %>% summarise(Mean_Return = mean(daily.returns)) %>% 

ggplot(aes(reorder(gics_sector, -Mean_Return), Mean_Return, fill = Mean_Return))+
  geom_col()+
  coord_flip()+
  theme_classic()+
  theme(legend.position="none")+
  labs(title = "Average Return by Sector Over Last Month", y = "Mean Return", x = "GICS Sector")
```

`



```{r}
plot = avg_return %>% ggplot(aes(avg_return, Volatility))+
  geom_text(aes(label = symbol), size = 3)+
  labs(title = "Average Return vs Volatility Over Last 3 Months In SP500", x = "Average Return", subtitle = "Data Source: Yahoo Finance")+
  theme_minimal()
    
plot 
```

```{r}
avg_return = avg_return %>% 
  mutate(Indicator = case_when(symbol %in% c('FOX', 'FOXA') ~ "Fox Corporation",
                               TRUE ~ "Rest of SP500"))
```


```{r}
plot = avg_return %>% ggplot(aes(text = paste("Average Return", avg_return), text = paste("Volatility", Volatility), color = Indicator, text = security))+
  geom_text(aes(x = avg_return, y = Volatility, label = symbol), size = 3)+
  labs(title = "Average Return vs Volatility Over Last 3 Months In SP500", x = "Average Return", subtitle = "Data Source: Yahoo Finance")+
  theme_minimal()

    
ggsave("sp500plot.png", plot = plot)
```


```{r}
p <- plot_ly(avg_return, x = ~avg_return, y = ~Volatility, type = 'scatter',
        mode = 'text', text = ~symbol, textposition = 'middle right',
        color = ~Indicator, colors = 'Set1',
        textfont = list(size = 10)) %>%
  layout(title = 'Average Return vs Volatility Over Last 3 Months In SP500', 
         xaxis = list(title = 'Averaage Return', zeroline = FALSE
                      ),
         yaxis = list(title = 'Volatility'
                      ))
p

```

```{r}
symbols = c('FOX', 'FOXA')
fox = tq_get(symbols, from = '2017/01/01')
fox %>% 
  ggplot(aes(date, adjusted))+
  geom_line()+
  geom_smooth(method = 'lm', alpha = 0.3)+
  facet_wrap(~symbol)+
  theme_classic()

  
  
```


```{r, fig.height=8}
plot + facet_wrap(~gics_sector, ncol = 3) 
```


```{r}
zoomed = tickers_df %>% 
  filter(symbol %in% c('FOX', 'FOXA')) %>% 
  ggplot(aes(date, adjusted))+
  geom_point(alpha = 0.3, color = 'black')+
  geom_line()+
  facet_wrap(~symbol)+
  theme_classic()
ggplotly(zoomed)
```

